---
title: "homework 2"
author: Purnima Sharma
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.


```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.

```{r}
precip_2018 =
  read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
     sheet = "2018 Precipitation",
     skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 =
  read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
     sheet = "2017 Precipitation",
     skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

precip_df =
  left_join(precip_df, month_df, by = "month")
```
This dataset contains information from the Mr Trashwheel trash collector in Baltimore, Maryland.AS trash enters the inner harbor, the trashwheel collects  that trash, and stores it in a dumspter. The dataset contains information on year, month and trash collected, includign some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data. In this dataset:

The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2

Read and clean the Transit data. Convert "Entry" variable from character to logical. 

```{r}
transit_df_raw = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE ))

transit_df = select(transit_df_raw, line:ada, -exit_only, -staffing, -staff_hours)
```

The Transit dataset consists of 1,868 observations and 32 variables, with information on subway entrance and exit locations in New York city, different routes available, stations, and ADA compliance information. The above code updates the dataset by keeping 19(use code for columns) of the 32 variables, by cleaning the names, and by converting "entry" variable from character to logical form, assigning true values to YES and false to NO. In this dataset:

There are `r distinct(transit_df, station_name, line)` distinct stations in the dataset. 
