---
title: "homework 2"
author: Purnima Sharma
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.


```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.

```{r}
precip_2018 =
  read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
     sheet = "2018 Precipitation",
     skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 =
  read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
     sheet = "2017 Precipitation",
     skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

precip_df =
  left_join(precip_df, month_df, by = "month")
```
This dataset contains information from the Mr Trashwheel trash collector in Baltimore, Maryland.AS trash enters the inner harbor, the trashwheel collects  that trash, and stores it in a dumspter. The dataset contains information on year, month and trash collected, includign some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data. In this dataset:

The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2

Read and clean the Transit data. Convert "Entry" variable from character to logical. 

```{r}
transit_df_raw = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE ))

transit_df = select(transit_df_raw, line:ada, -exit_only, -staffing, -staff_hours) 
```

The Transit dataset consists of 1,868 observations and 32 variables, with information on subway entrance and exit locations in New York city, different routes available, stations, and ADA compliance information. The above code updates the dataset by keeping `r ncol(transit_df)` of the 32 variables, by cleaning the names, and by converting "entry" variable from character to logical form, assigning true values to YES and false to NO. In this dataset:

There are `r nrow(distinct(transit_df, station_name, line))` distinct stations in the dataset. 

There are `r transit_df %>% filter(ada == TRUE) %>% distinct(station_name, line) %>% nrow()` ADA compliant stations. 

Proportion of station entrances/exits without vending that allow entrance is  `r transit_df %>% filter(entry == TRUE & vending == "NO") %>% nrow()` out of  `r transit_df %>% filter(vending == "NO") %>% nrow()`.


Create route name and route number variables.

```{r}
transit_df_tidy = 
  transit_df %>% 
  mutate_at(vars(route1:route11), as.character,
   replace_na, 'None') %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  ) 
```

There are `r transit_df_tidy %>% filter(route_number == 'A') %>% distinct(station_name, line) %>% nrow()` distinct stations that serve the A train.

 `r transit_df_tidy %>% filter(route_number == 'A'& ada == TRUE) %>% distinct(station_name, line) %>% nrow()` stations serving A train are ADA compliant.


## Problem 3

Import and clean 'pos-month', 'unemployment' and 'snp' data from FiveThirtyEight dataset.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  ) 
month_df = mutate(month_df, month = as.character(month))

pols_df = 
   read_csv("./data/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(month = as.character(month))
  
 pols_df =
   left_join(pols_df, month_df, by = "month") %>% 
   mutate(month = month_name) %>% 
   mutate(president = 
            case_when(
              prez_dem == 1 ~ "dem",
              prez_gop == 1 ~ "gop")) %>% 
   select(-month_name, -day, -prez_dem, -prez_gop)
 
 snp_df = 
   read_csv("./data/snp.csv") %>% 
  separate(date, into = c("month", "day", "year")) 
 snp_df =
    left_join(snp_df, month_df, by = "month") %>% 
   mutate(month = month_name) %>% 
   select(-month_name, -day) %>% 
   mutate(year = as.integer(year)) %>% 
   relocate(year, month)
 
 unemployment_df =
    read_csv("./data/unemployment.csv") %>% 
   pivot_longer(
      Jan:Dec,
    names_to = "month",
    values_to = "unemployed"
   ) %>% 
  mutate(month = recode(month, "Jan" = "January", "Feb" = "February", "Mar" = "March", "Apr" = "April", "Jun" = "June", "Jul" = "July", "Aug" = "August", "Sep" = "September", "Oct" = "October", "Nov" = "November", "Dec" = "December"))
```

Join dataset "snp" into "pols", and "unemployment" into the result.

```{r}
snp_pols_df = 
  left_join(pols_df, snp_df, by = c("year", "month"))

final_df =
  left_join(snp_pols_df, unemployment_df, by = c("year", "month"))
```


The three datasets used in this problem are from the website FiveThirtyEight which focuses on poll analyses in politics, science, and many other fields. The datasets pols_month, snp, and unemployment all run through the similar timeframe, from the year 1947 to 2015. The data in pols_month primarily depicts the number of politicians who are democratic or republican at any given time. The snp-data is about the S&P stock market at any given date of observation, and the unemployment-data shows the percent unemployed for a given month in a given year in the timeframe. 
The resulting dataset obtained by merging snp and then unemplyment into pols_month was obtained using year and month as keys, and contains `r nrow(final_df)` rows and `r ncol(final_df)` columns. It shows monthly information from `r summarise(final_df, min(year))`to `r summarise(final_df, max(year))` on the party affiliation of the president, the stock market, and the unemployment rate.

